---
title: "Homework 2 Report"
output:
  html_document: default
  pdf_document: default
---


## Name

Please enter your name here.

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
dataset=read.csv("maacs.csv.gz")
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here
check_fev1_age_relationship <- function(dataset) {
  sorted_data <- dataset[order(dataset$age), ]
  for (i in 2:nrow(sorted_data)) {
    if (sorted_data$fev1[i] <= sorted_data$fev1[i - 1]) {
      return(FALSE)  
    }
  }
  return(TRUE)
}

dataset_clean <- dataset[!is.na(dataset$age) & !is.na(dataset$fev1), ]
result <- check_fev1_age_relationship(dataset_clean)
print(result)
```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}
## Add your code here
model <- lm(fev1 ~ age, data = dataset)
summary(model)
```


Write your data analysis statement interpreting the regression model here:
It is estimated that FEV1 increases by around 0.171 liters for each increase in age by 1. 


## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results.

NOTE: At least one supporting premise should use a plot.

```{r, warning=F}
## Add your code here
correlation <- cor(dataset$age, dataset$fev1, use = "complete.obs")
print(correlation)

model <- lm(fev1 ~ age, data = dataset)
summary(model)

library(ggplot2)
ggplot(dataset, aes(x = age, y = fev1)) +
  geom_point() +  # Scatter plot without the regression line
  labs(title = "Scatter Plot of Age vs FEV1",
       x = "Age (Years)",
       y = "FEV1") +
  theme_minimal()
```


Write the three supporting premise statements here:

1. There is a strong positive correlation between FEV1 and age.

2. The slope for the linear regression model is positive. 

3. The scatterplot with the 2 variables shows an upward trend 


## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Function for supporting premise statement 1
check_correlation <- function(dataset) {
  correlation <- cor(dataset$age, dataset$fev1, use = "complete.obs")
  if (correlation > 0.7) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```


```{r}
## Function for supporting premise statement 2
check_slope_positive <- function(dataset) {
  model <- lm(fev1 ~ age, data = dataset)
  if (summary(model)$coefficients[2, 1] > 0) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```


```{r}
## Function for supporting premise statement 3
check_scatterplot_upward_trend <- function(dataset) {
  plot <- ggplot(dataset, aes(x = age, y = fev1)) +
    geom_point() +  
    labs(title = "Scatter Plot of Age vs FEV1",
         x = "Age (Years)",
         y = "FEV1") +
    theme_minimal()
  print(plot)
  
  # Simulate hypothetical data with an upward trend
  set.seed(42)  # for reproducibility
  age_sim <- seq(0, 20, length.out = 100)  # 100 age values from 0 to 20
  fev1_sim <- 0.8 + 0.1 * age_sim + rnorm(100, mean = 0, sd = 0.2)  # linear trend + noise
  
  simulated_data <- data.frame(age = age_sim, fev1 = fev1_sim)
  
  # Create the hypothetical plot
  hypothetical_plot <- ggplot(simulated_data, aes(x = age, y = fev1)) +
    geom_point(color = "blue", alpha = 0.6) +  # simulated points
    geom_smooth(method = "lm", se = FALSE, color = "red") +  # regression line
    labs(title = "Hypothetical Scatter Plot with Upward Trend",
         x = "Age (Years)",
         y = "FEV1") +
    theme_minimal()
  
  # Display the plot
  print(hypothetical_plot)
}
```

Execute each of your function and show that the produce the expected output.

```{r, message=F}
result1 <- check_correlation(dataset_clean)
print(result1)
result2 <-check_slope_positive(dataset_clean)
print(result2)
check_scatterplot_upward_trend(dataset_clean)
```



## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.


